Food-web structure, stochasticity and response diversity drive temporal community stability in complex communities

### 2023-12-12 _targets.Rmd
New SEM figure and prettier method figure

      "ct_alive" = .1,

@@

      "w_avg_tlvl" = 1,

@@

```{r}

@@

```{r tar-make}
tar_make()
tar_make_future(workers = 3)
```

@@

@@

```{r plot-sim}
load(here("data", "sim_for_plot"))
length(unique(sim_for_plot$sim_id))
sim_for_plot %<>%
  left_join(
    tar_read(sim_fw)
  )
stoch <- 0.3
hill <- 2
env_corr <- 0
unique(sim_for_plot$env_stoch)
```

```{r}
sim_for_plot %>%
  filter()

@@

```{r}
sim_for_plot$int_strength[[1]] %>% dim
p_low_z_low_rich <- plot_sim(x = sim_for_plot,
  id = low_z_low_rich$sim_id[[1]],
  col_palette = NULL,
  node_size_scale = .3,
  tmax = 200
)
p_low_z_high_rich <- plot_sim(x = sim_for_plot,
  id = low_z_high_rich$sim_id[[2]],
  col_palette = NULL,
  node_size_scale = .5
)
p_high_z_low_rich <- plot_sim(
  x = sim_for_plot,
  id = high_z_low_rich$sim_id[[5]],
  col_palette = NULL,
  node_size_scale = .3)
p_high_z_high_rich <- plot_sim(
  x = sim_for_plot,
  id = high_z_high_rich$sim_id[[3]],
  col_palette = NULL,
  node_size_scale = .4
)

@@

high_resp_div_low_rich <- sim_for_plot %>%
  filter(fw_id == fwid, Z == z, env_stoch == 0.6, rho == 0)
  #filter(env_stoch == .6, Z == 1, S == 10, rho == 0) %>%
  #arrange(desc(async))
p_high_resp_div_low_rich <- plot_sim(
  x = sim_for_plot,
  id = high_resp_div_low_rich$sim_id[[1]],
  col_palette = NULL,
  node_size_scale = .2,
  title = "High response diversity"
)

low_resp_div_low_rich <- sim_for_plot %>%
  filter(fw_id == fwid, Z == z, env_stoch == 0.6, rho == 1) %>%
  arrange(desc(async))
p_low_resp_div_low_rich <- plot_sim(
  x = sim_for_plot,
  id = low_resp_div_low_rich$sim_id[[1]],
  col_palette = NULL,
  node_size_scale = .2,
  title = "Low response diversity"
)

>>>>

      "ct_alive" = .05,

@@

      "w_avg_tlvl" = 1.0,

@@

## Partial residuals SEM

```{targets partial_sem_relations}
list(
  tar_target(scaled_sim_fw_sem,
    sim_fw_sem %>%
      mutate(across(where(is.numeric), ~scale(.x)[, 1]))
    ),
  tar_target(scaled_sem_simplified_model_list,
    map(sem_simplified_model_list,
      ~update(.x, data = scaled_sim_fw_sem)
      )),
  tar_target(scaled_sem_simplified,
    as.psem(scaled_sem_simplified_model_list)
    ),
  tar_target(formula_sem_bivariate_relations,
    c(
      "pop_stab ~ w_avg_tlvl",
      "pop_stab ~ h",
      "pop_stab ~ env_stoch",
      "sae_total ~ richness",
      "sae_total ~ ct_alive",
      "sae_total ~ w_avg_tlvl",
      "sae_total ~ avg_int_strength",
      "cpe ~ avg_int_strength",
      "cpe ~ resp_div"
    )
    )#,
  #tar_target(partial_resid_sem_simplified, {
  #  partialResid(as.formula(formula_sem_bivariate_relations),
  #          modelList = scaled_sem_simplified,
  #          data = scaled_sim_fw_sem
  #        )},
  #        pattern = map(formula_sem_bivariate_relations)
  #),
  #tar_target(sem_simplified_bivariate_relations,
  #  tibble(sem_formula = formula_sem_bivariate_relations) %>%
  #    separate(
  #      col = sem_formula,
  #      into = c("response", "predictor"),
  #      sep = " ~ ", remove = FALSE) %>%
  #  mutate(
  #    data = partial_resid_sem_simplified,
  #    coefs = map2(response, predictor,
  #      ~coefficients(scaled_sem_simplified[[.x]])[c("(Intercept)", .y)]
  #      ),
    #  plot = pmap(list(
    #      data = data,
    #      response = response,
    #      predictor = predictor,
    #      coef = coefs
    #      ),
    #    function (data, response, predictor, coef, ext_data) {
    #      if (predictor %in% c("h", "env_stoch", "resp_div")) {
    #        p <- data %>%
    #          ggplot(aes(
    #              y = yresid, x = xresid,
    #              group = cut(round(xresid, 1),
    #                breaks = length(unique(ext_data[[predictor]]))
    #              )
    #              )) +
    #          geom_boxplot(alpha = .1)
    #      } else {
    #        p <- data %>%
    #          ggplot(aes(y = yresid, x = xresid)) +
    #          geom_point(alpha = .1)
    #      }
    #      p +
    #        geom_abline(
    #          intercept = coef["(Intercept)"],
    #          slope = coef[2],
    #          linewidth = 1
    #          ) +
    #        labs(
    #          x = var_replacement()[predictor],
    #          y = var_replacement()[response]
    #          ) +
    #        coord_cartesian(expand = FALSE) +
    #        theme_half_open()

    #    }, ext_data = scaled_sim_fw_sem)
    #)
  #)
)
```

@@

```{r tar-make}
#tar_make()
tar_make_future(workers = 3)
```

@@

```{r}
tar_load(c(formula_sem_bivariate_relations, scaled_sem_simplified, scaled_sim_fw_sem))
sem_simplified_bivariate_relations <- tibble(formula = formula_sem_bivariate_relations) %>%
      separate(
        col = formula,
        into = c("response", "predictor"),
        sep = " ~ ", remove = FALSE) %>%
    mutate(
      data = map(formula_sem_bivariate_relations,
        ~partialResid(as.formula(.x),
            modelList = scaled_sem_simplified,
            data = scaled_sim_fw_sem
          )
        ),
      coefs = map2(response, predictor,
        ~coefficients(scaled_sem_simplified[[.x]])[c("(Intercept)", .y)]
        ),
      plot = pmap(list(
          data = data,
          response = response,
          predictor = predictor,
          coef = coefs
          ),
        function (data, response, predictor, coef, ext_data) {
          if (predictor %in% c("h", "env_stoch", "resp_div")) {
            p <- data %>%
              ggplot(aes(
                  y = yresid, x = xresid,
                  group = cut(round(xresid, 1),
                    breaks = length(unique(ext_data[[predictor]]))
                  )
                  )) +
              geom_boxplot(alpha = .1)
          } else {
            p <- data %>%
              ggplot(aes(y = yresid, x = xresid)) +
              geom_point(alpha = .1)
          }
          p +
            geom_abline(
              intercept = coef["(Intercept)"],
              slope = coef[2],
              linewidth = 1
              ) +
            labs(
              x = var_replacement()[predictor],
              y = var_replacement()[response]
              ) +
            coord_cartesian(expand = FALSE) +
            theme_half_open()

        }, ext_data = scaled_sim_fw_sem)
    )
```

```{r}
label_sem <- as.character(as.roman(seq_len(nrow(sem_simplified_bivariate_relations))))
p_sem_simplified_bivariate_relations <-
  plot_grid(
    plotlist = sem_simplified_bivariate_relations$plot,
    label_x = 1, hjust = 1.0,
    labels = label_sem
  )
sem_file <- here("report", "figures", "sem_stab_simplified.pdf")
p_sem <- ggdraw() +
  draw_image(magick::image_read_pdf(sem_file, density = 600),
    x = 0, y = 0,
    width = 1, height = 1,
    hjust = 0, halign = 0, valign = 0
  )
p_sem_tot <- plot_grid(
  p_sem, p_sem_simplified_bivariate_relations,
  nrow = 2, rel_heights = c(.6, 1),
  labels = "auto"
)
ggsave_multiple(
  paste0("p_sem_tot", c(".png", ".pdf")),
  plot = p_sem_tot,
  path = here::here("report", "figures"),
  scale = 2.4,
  dpi = 400,
  width = 80,
  height = 80 * 1.6,
  units = "mm")
```

```{r}
p_sem_fig <- plot_grid(
  p_sem, p_sem_simplified_bivariate_relations,
  print(
    p_sem_tot_stab +
      guides(
        color = guide_legend(override.aes = list(linetype = 0, alpha = 1))
      ) +
      theme(
        legend.title = element_blank(),
        legend.position = c(.005, .995),
        legend.justification = c("left", "top"),

      )
    ),
  nrow = 3, rel_heights = c(.7, 1, .6),
  labels = "auto"
)
ggsave_multiple(
  paste0("p_sem_fig", c(".png", ".pdf")),
  plot = p_sem_fig,
  path = here::here("report", "figures"),
  scale = 2.4,
  dpi = 400,
  width = 80,
  height = 80 * 2.0,
  units = "mm")
```

@@

```{r plot-sim}
load(here("data", "sim_for_plot"))
length(unique(sim_for_plot$sim_id))
sim_for_plot %<>%
  left_join(
    tar_read(sim_fw)
  )
stoch <- 0.3
hill <- 2
env_corr <- 0
label_var <- c("w_avg_tlvl", "stab_com", "pop_stab", "async", "sae_total", "cpe")
unique(sim_for_plot$env_stoch)
```

@@

```{r}
sim_for_plot$int_strength[[1]] %>% dim
p_low_z_low_rich <- plot_sim(x = sim_for_plot,
  id = low_z_low_rich$sim_id[[1]],
  col_palette = NULL,
  node_size_scale = .3,
  tmax = 200,
  var_label = label_var
)
p_low_z_high_rich <- plot_sim(x = sim_for_plot,
  id = low_z_high_rich$sim_id[[2]],
  col_palette = NULL,
  node_size_scale = .5,
  var_label = label_var
)
p_high_z_low_rich <- plot_sim(
  x = sim_for_plot,
  id = high_z_low_rich$sim_id[[5]],
  col_palette = NULL,
  node_size_scale = .3,
  var_label = label_var)
p_high_z_high_rich <- plot_sim(
  x = sim_for_plot,
  id = high_z_high_rich$sim_id[[3]],
  col_palette = NULL,
  node_size_scale = .4,
  var_label = label_var
)

@@

high_resp_div_low_rich <- sim_for_plot %>%
  filter(fw_id == fwid, Z == z, env_stoch == 0.6, rho == 0)
  #filter(env_stoch == .6, Z == 1, S == 10, rho == 0) %>%
  #arrange(desc(async))
p_high_resp_div_low_rich <- plot_sim(
  x = sim_for_plot,
  id = high_resp_div_low_rich$sim_id[[1]],
  col_palette = NULL,
  node_size_scale = .2,
  title = "High response diversity",
  var_label = label_var
)

low_resp_div_low_rich <- sim_for_plot %>%
  filter(fw_id == fwid, Z == z, env_stoch == 0.6, rho == 1) %>%
  arrange(desc(async))
p_low_resp_div_low_rich <- plot_sim(
  x = sim_for_plot,
  id = low_resp_div_low_rich$sim_id[[1]],
  col_palette = NULL,
  node_size_scale = .2,
  title = "Low response diversity",
  var_label = label_var
)

### 2023-12-12 paper/manuscript.Rmd
New SEM figure and prettier method figure

```{r}
sem_cap <- paste0(
  "Structural Equation Model linking initial model parameters to resulting
  temporal stability of biomass. Arrows display positive (continuous) and
  negative (dashed) standardised effects. Initial parameters were initial
  species richness, initial connectance, Prey-Predator Mass Ratio (PPMR), hill
  exponent, variance of stochastic mortality rates (Env. stochasticity), and
  species response diversity. We only
  displayed standardised coefficients whose absolute values are superior to 0.1
  are displayed. A more complete SEM containing further partition of Statistical
  Averaging Effect and Compensatory effects is displayed in Fig. SXX.")
```

```{r sem, fig.cap=sem_cap, fig.id = "sem"}
knitr::include_graphics(here("report", "figures", "sem_stab_simplified.pdf"))
```

>>>>

```{r}
sem_cap <- paste0(
  "Structural Equation Model linking initial model parameters to resulting
  temporal stability of biomass. (a) Arrows display positive (continuous) and
  negative (dashed) standardised effects. Initial parameters were initial
  species richness, initial connectance, Prey-Predator Mass Ratio (PPMR), hill
  exponent, variance of stochastic mortality rates (Env. stochasticity), and
  species response diversity. Only the standardised coefficients whose absolute
  values are superior to 0.1 are displayed. A more complete SEM containing
  further partition of Statistical Averaging Effect and Compensatory effects is
  displayed in Fig. SXX. (b) Partial relationships derived from the SEM, points
  displaying the partial standardized residuals.")
```

```{r sem, fig.cap=sem_cap, fig.id = "sem"}
knitr::include_graphics(here("report", "figures", "p_sem_tot.pdf"))
```
